---
title: "Data Visualization of Student Performance"
author: "Matthew Flanders"
date: "1/30/2022"
output: pdf_document
---

```{r, include=F }
library( tidyverse)
library(Sta486Project1mtf83)
```

\newpage
```{r,echo=FALSE}

data( "mathData" )

ggplot( mathData, aes( x = G3, y=..density..))+
  geom_histogram( bins = 10)+
  facet_grid( . ~ Walc)+
  geom_density()+
  labs(title = "Final grade by increased weekend alcohol consumption")+
  xlab("Final Grade")
```
**Figure 1.** I felt that this graph had interesting behavior when the 2( low alcohol consumption) group had a high density of low test scores. I feel that there may be under reporting, or some bias to the survey results where students are under estimating thier alcohol consumption and that may have a negative effect on their studies.  
\newpage
```{r,echo=FALSE}

ggplot( mathData, aes( x = G3))+
  geom_histogram( bins = 10)+
  facet_grid( . ~ nursery)+
  aes(y=..density..)+
  geom_density()+
  ggtitle("Final Grade by Nursery school Attandance")+
  xlab("Final Grade")

```
**Figure 2.** This graph is somewhat confirming my assumptions on that a child getting a headstart on thier education can benifit them in the long run. 
\newpage

```{r,echo=FALSE}
data( "porData")

ggplot( porData, aes( x = G3, y=..density..))+
  geom_histogram( bins = 10)+
  facet_grid( . ~ school)+
  geom_density()+
  ggtitle("Final grade by School")+
  xlab("Final Grade")
```
**Figure 3.** Here I found the differences in the density plot interesting for the two schools. I would like to some further analysis on some of the school demographics to see what may be causing the differences seen.
\newpage
```{r,echo=FALSE}
ggplot( porData, aes( x = G3, y=..density.., color = school))+
  geom_density()
```
**Figure 4.** By stacking the two density plots I felt that the differences were much more apparent. 
\newpage
```{r,echo=FALSE}

ggplot( porData, aes( x = G3, y=..density..))+
  geom_histogram( bins = 10)+
  facet_grid( . ~ schoolsup)+
  geom_density()+
  ggtitle("Final grade by School Support for Students")+
  xlab("Final Grade")

```
**Figure 5.** I think that this graph was interesting in that there were many more students performing poorly with school support, and I also find the large gap between peaks interesting. It could be that some students are able to really benefit from help while some may not be interested in school.
\newpage
```{r,echo=FALSE}

ggplot( porData, aes( x = G3))+
  geom_histogram( )+
  facet_grid( . ~ nursery)+
  aes(y=..density..)+
  geom_density()+
  ggtitle("Final grade by Nursery school Attandance")+
  xlab("Final Grade")

```
**Figure 6.** Another view of the results of attending nursery school. 




```{r}

ggplot( mathData, aes( x = school, y = G3 )) + geom_boxplot()
m1 <- lm( G3 ~ Walc, mathData)
anova(m1)

```
